```{r setup, echo=FALSE}
opts_knit$set(upload.fun = WrapWordpressUpload, base.url = NULL) 
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE)
````

Today, I'm going to finally dip my toe in the water of something I've wanted to do for a long time. I'd wanted to do this with Michael Caine film appearances- and I may yet- but tonight, it'll be a bit simpler. I'm going to draw a network graph, showing the relationship between band members of the Grateful Dead and a couple of their solo and non-Dead projects. Information is from my head or wikipedia, reflecting the most recent band lineups. 

First, I'll create data frames for several bands.

```{r BuildBands, echo=TRUE}
dfGratefulDead = data.frame(Musician = c("Jerry Garcia", "Bob Weir", "Phil Lesh"
                                          , "Bill Kreutzmann", "Mickey Hart", "Pigpen")
                            , Instrument = c("guitar", "guitar", "bass"
                                           , "drums", "drums", "keyboards"))

dfJerryGarciaBand = data.frame(Musician = c("Jerry Garcia", "John Kahn", "Melvin Seals"
                                             , "David Kemper", "Jaclyn LaBranch", "Gloria Jones")
                               , Instrument = c("guitar", "bass", "keyboards"
                                              , "drums", "vocals", "vocals"))
```

```{r, echo=FALSE}
dfRatDog = data.frame(Musician = c("Bob Weir", "Jay Lane", "Robin Sylvester"
                                    , "Jeff Chimenti", "Kenny Brooks", "Mark Karan")
                      , Instrument = c("guitar", "drums", "bass"
                                       , "keyboards", "saxophone", "vocals"))

dfOtherOnes = data.frame(Musician = c("Bob Weir", "Phil Lesh", "Mickey Hart"
                                       , "Bill Kreutzmann", "Jimmy Herring", "Rob Barraco"
                                       , "Jeff Chimenti", "Joan Osborne")
                         , Instrument = c("guitar", "bass", "drums"
                                          , "drums", "guitar", "keyboards"
                                          , "keyboards", "vocals"))
library(xtable)
```
**Grateful Dead**
```{r, echo=FALSE, results = 'asis'}
print(xtable(dfGratefulDead), type = "html")
```
**Jerry Garcia Band**
```{r, echo = FALSE, results = 'asis'}
print(xtable(dfJerryGarciaBand), type = "html")
```

I've created data frames for RatDog and the Other Ones as well, but the code and results are not shown. Next, I'll code a helper function to create relationships between band members.
```{r ReadData, warning = FALSE, message = FALSE, fig.width = 10}
CreateRelation = function(dfBand, BandName){

  indices = combn(length(dfBand$Musician), 2)
  
  dfRelate = data.frame(from = dfBand$Musician[indices[1, ]]
                        , to = dfBand$Musician[indices[2, ]])
  dfRelate$Band = BandName
  
  dfRelate
  
}
```

With that in place, I can create relationships and then rbind the results. I tried adding a color column for a musicians instrument, but haven't had any luck in being able to apply it in the plot. Suggestions are welcome.

```{r, echo=TRUE}
dfGDrelate = CreateRelation(dfGratefulDead, "GratefulDead")
dfJGBrelate = CreateRelation(dfJerryGarciaBand, "Jerry Garcia Band")
dfRatDogRelate = CreateRelation(dfRatDog, "RatDog")
dfOtherOnesRelate = CreateRelation(dfOtherOnes, "The Other Ones")

dfMusicians = rbind(dfGratefulDead, dfJerryGarciaBand, dfRatDog, dfOtherOnes)

dfMusicians = dfMusicians[!duplicated(dfMusicians), ]

dfInstrumentColor = data.frame(Instrument = c("bass", "drums", "guitar", "keyboards", "vocals", "saxophone")
                               , Color = c("black", "beige", "red", "blue", "green", "yellow"))
dfMusicians = merge(dfMusicians, dfInstrumentColor)
dfMusicians = dfMusicians[, c("Musician", "Instrument", "Color")]

dfRelations = rbind(dfGDrelate, dfJGBrelate, dfRatDogRelate, dfOtherOnesRelate)
```

```{r fig.width=10, fig.height=10}
library(igraph)

g = graph.data.frame(dfRelations, directed=FALSE, vertices=dfMusicians)

set.seed(1234)
plot(g, vertex.color = g$Color)
```

Note the mashup of Bobby, Mickey, Bill and Phil in the center. They've had a lot of crossover post 1995. If I didn't have to work tomorrow, I'd probably spend the next few hours playing with this. Developing a chart of all the permutations of Yes, Asia, King Crimson, etc. could take the better part of an evening.

```{r SessionInfo}
sessionInfo()
```