```{r setup, echo=FALSE}
opts_knit$set(upload.fun = WrapWordpressUpload, base.url = NULL) 
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE)
````
I've finally had some success at munging some HTML. For quite some time, I've wanted to render a county level choropleth for US presidential election results. The numbers are all there on [Politico.com](www.politico.com), but attempts to use readHTMLTable never returned the full set of data. It still doesn't, but I have sorted out how to get all of the results I want. It takes a fair bit of work, but- once the smoke clears- doesn't seem too crazy.

First, we'll fetch some raw HTML for North Carolina.
```{r FetchHTMLfile}
library(XML)

URL = "http://www.politico.com/2012-election/results/president/north-carolina/"
content.raw = htmlParse(URL, useInternalNodes = TRUE)
```

Inspection of the tables which get returned tell us that the second element in the list has the data we need. Attempts to extract the information lead us to take a slightly different approach. First, we'll get all the nodes with a "tbody" element. Each of these nodes may be treated as a table.

```{r GetCountyTables}
tables <- getNodeSet(content.raw, "//table") 
counties = getNodeSet(tables[[2]], "//tbody")
counties = counties[-1]

countyTables = lapply(counties, readHTMLTable, header = FALSE, stringsAsFactors = FALSE)
```

The table we get isn't quite what we want.
```{r ShowWrongTable}
head(countyTables[[1]])
```

A couple helper functions will fetch the county name and move the cells to a sensible location.
```{r HelperFunctions}
GetCountyName = function(dfCounty){
  strCounty = dfCounty[1,1]
  strCounty = strsplit(strCounty, " ")
  strCounty[[1]][1]
}

MungeTable = function(dfCounty){
  
  if (ncol(dfCounty) != 5 ) return (data.frame())
  
  dfCounty[1, 1]= GetCountyName(dfCounty)

  dfCounty[-1, 2:5] = dfCounty[-1, 1:4]
  
  dfCounty[, 1] = dfCounty[1, 1]
  
  colnames(dfCounty) = c('CountyName', 'Candidate', 'Party', 'Pct', 'Votes')
  
  dfCounty$Votes = gsub(",", "", dfCounty$Votes)
  dfCounty$Votes = as.numeric(dfCounty$Votes)
  
  dfCounty$Pct = NULL
  
  dfCounty
}

correctTable = MungeTable(countyTables[[1]])
head(correctTable)
```

With that done, it's a simple thing to munge each data frame and then bind the results into a single data frame. 
```{r MungeCounties}
counties = lapply(countyTables, MungeTable)
dfNorthCarolina = do.call("rbind", counties)

```

A plot shows that Obama won in counties with a high population, but didn't do as well in smaller counties. I'll draw some better charts tomorrow.
```{r PlotResults, , fig.width = 10}
library(ggplot2)
ggplot(dfNorthCarolina, aes(x = CountyName, y = Votes, fill = Party)) + geom_bar(stat = "identity")
```

This required getting very, very familiar with the underlying HTML structure. That's a hassle, but hardly impossible. Tomorrow, this will become a map and I'll make some inferences about voting patterns and demographics.
```{r SessionInfo}
sessionInfo()
```