```{r setup, echo=FALSE}
opts_knit$set(upload.fun = WrapWordpressUpload, base.url = NULL) 
opts_chunk$set(fig.width=10, fig.height=10, cache=FALSE, echo=TRUE, warning=FALSE, message=FALSE, error=FALSE)
````
I like to use [Goodreads](www.goodreads.com) to keep track of which books I'm reading (and not reading). They very helpfully sent me an e-mail to inform me how many books I've read so far in 2013. The number is 19. Hardly an impressive number, but between job, family and trying to develop my R skills, I'm not embarassed by that either. However, it's not as though I'm reading James Joyce. This year reveals a great interest in the works of [Robert B. Parker.](https://www.goodreads.com/author/show/397.Robert_B_Parker). There are a few omissions as well- all my fault. I never bothered to enter [Naked Statistics](https://www.goodreads.com/book/show/13707560-naked-statistics?from_search=true) and [Medium Raw](https://www.goodreads.com/book/show/7324659-medium-raw?from_search=true) was never listed as having been finished. That understood, let's see if I can learn anything about my reading habits.

I've not yet come to grips with the Goodreads API. It was the work of only a few minutes to copy and paste my collection of books into [Excel](http://www.r-bloggers.com/excel-fanaticism-and-r/) and strip out the nonsense to produce a quick CSV file. I'll load that in and have a look.
```{r ReadCSV}
dfBooks = read.csv("./Data/Goodreads.csv", stringsAsFactors = FALSE)
colnames(dfBooks)[6] = "DateRead"
colnames(dfBooks)[7] = "DateAdded"
dfBooks$DateAdded = gsub("[edit]", "", dfBooks$DateAdded)
dfBooks$DateAdded = gsub(" ", "", dfBooks$DateAdded)
dfBooks$DateAdded = as.Date(dfBooks$DateAdded, "%m/%d/%Y")
``` 

When did I add books to my list? Most of the time this will be one book on a particular day, but there are exceptions. I'll look at this timeline in a few different ways.

```{r Explore}
dfAgg = aggregate(x = dfBooks$DateAdded, by = list(dfBooks$DateAdded), FUN = length)
colnames(dfAgg) = c("DateAdded", "NumBooks")
plot(x = dfAgg$DateAdded, y = dfAgg$NumBooks, pch = 19, xlab="Date", ylab="Number of books")
```

That giant bump is when GoodReads switched over from their Facebook app. I'll zero in on something relevant.

```{r Recent}
dfAgg = subset(dfAgg, DateAdded >= as.Date("2012-01-01"))
plot(x = dfAgg$DateAdded, y = dfAgg$NumBooks, pch = 19, xlab="Date", ylab="Number of books")
```

There's another crazy blip near the start of 2012. I've no idea why everything defaults to sometime in the middle of last year. I've been reading books for quite a long time and there's not a lot of variation. 

No matter. This isn't terribly informative. I'm going to try looking at this differently. The date I finished the book isn't often available and is formatted in such a way that it resists easy munging. I'm going alter it manually for the 44 books that I started in the last three months of 2012.

```{r Timeline}
dfBooks2 = read.csv("./Data/Goodreads2.csv", stringsAsFactors = FALSE)
dfBooks2$DateRead = as.Date(dfBooks2$DateRead, "%m/%d/%Y")
dfBooks2$DateAdded = as.Date(dfBooks2$DateAdded, "%m/%d/%Y")

library(timeline)
# I need to sample a random integer so that we have a unique identifier for the book
# timeline's default is to use the book title as the label.
set.seed(1234)
dfBooks2$GroupCol = sample(nrow(dfBooks2))
timeline(dfBooks2[!is.na(dfBooks2$DateRead), ], group.col = "GroupCol", start.col = "DateAdded", end.col = "DateRead", label.col = "title")
```

Cool. For the record, it didn't take me that long to read American Splendor. I think that's one where I had put the book aside for quite some time (it's a lot of short vignettes) and then finally marked it as read. The Rise and Fall of Prussia seems about right. I recall having read that in a couple concentrated bursts that were spaced far apart. I'm also struck by the fact that it took only a couple days more to read a scholarly account of the life of Jesus than it does for me to read a Spenser novel.

Tomorrow: Probably heteroskedasticity.

```{r SessionInfo}
sessionInfo
```