```{r setup, echo=FALSE}
opts_knit$set(upload.fun = WrapWordpressUpload, base.url = NULL) 
opts_chunk$set(fig.width=10, fig.height=10, cache=FALSE, echo=TRUE, warning=FALSE, message=FALSE, error=FALSE)
````

Penultimate post, I'm going to take a quick look at the Gini indicator for wealth inequality. Data comes from the [World Bank](http://data.worldbank.org/indicator/SI.POV.GINI).

I've downloaded the zipped file, decompressed it and given it a different name. I'm going to 

```{r ReadCSV}
dfGini = read.csv("./Data/Gini.csv", stringsAsFactors = FALSE, skip = 2)
colnames(dfGini) = gsub("X", "", colnames(dfGini))
library(reshape2)
mdf = melt(dfGini, id.vars = colnames(dfGini)[1:4])
mdf = mdf[!is.na(mdf$value), ]
colnames(mdf)[1:2] = c("Name", "Code")
library(plyr)
dfCount = ddply(mdf, "Code", .fun = nrow)
atLeast5 = dfCount$Code[dfCount$V1 >= 5]
mdf = mdf[mdf$Code %in% atLeast5, ]
``` 

This will give us a decent set of data. How does this look when we plot it?

```{r Plot}
library(ggplot2)
ggplot(mdf, aes(x = variable, y = value, group = Code)) + geom_line()
```

Ugh. That looks like nothing. It's a bit late and I'm not all that keen for insight. What countries have had the biggest reduction in income inequality?
```{r BiggestMovers}
dfMove = ddply(mdf, "Code", summarize, Diff = max(value) - min(value))
dfMove = dfMove[order(dfMove$Diff, decreasing = TRUE), ]
bigMoves = dfMove$Code[1:5]

ggplot(mdf[mdf$Code %in% bigMoves, ], aes(x = variable, y = value, group = Code)) + geom_line()
```

That's better, but still not very informative. I'm sleepy and going to bed.

```{r SessionInfo}
sessionInfo
```